Article Title

### 2013-12-17 plos.tex
further progress on Discussion.

\section*{Introduction}
\subsection*{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Sompolinsky:2001,Wilke:2002,Averbeck:2006,Josic:2009,Berens:2011,Ecker:2011}. In addition, neural correlations are hypothesized to reflect key features of functional connectivity in neural circuits.  Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.

@@

To illustrate challenges of estimating the correlation matrix from a finite sample, we briefly consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}. In covariance selection, the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero. We estimated the covariance matrix of a period of a somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (Fig.\;\ref{fig:01}\,A, B) and covariance selection (Fig.\;\ref{fig:01}\,C). Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales. The apparent similarity of the two estimates belies the utterly different partial correlation structure (Fig.\;\ref{fig:01}\,D, E): the regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remainder. Estimation by covariance selection is attractive for several reasons: First, it has fewer free parameters and is therefore less susceptible to sampling noise, yielding estimates that are, on average, closer to truth $\Sigma$. Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

@@

\section*{Discussion}
\subsection*{Functional connectivity from big neural data}
Transformational breakthroughs in neuroscience are anticipated to follow technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far. It far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques. Most common analyses of the functional connectivity rely on analyzing pairwise correlation coefficients, t-tests, ANCOVA, principal component analysis of correlated activity, and cross-correlations.  Traditionally, functional connectivity between regions has been defined as the level of correlation or synchronization between regions or cells. Without targeted stimulation and intracellular signals, synaptic connectivity could be inferred with some degree of confidence from the presence of sharp peaks or troughs in the cross-correlograms at offsets corresponding to typical synaptic transmission delays. However, even such effects can be explained by multiple effects besides monosynaptic connections while some strong synaptic connections do not necessarily produce a clear spike in the cross-correlogram. The situations is a bit more complicated in recording modalities, such as calcium imaging, in which only firing rates at lower temporal resolutions can be inferred and cross-correlograms with millisecond precision cannot be obtained. Therefore, some researchers have resorted to using simple correlations as proxies of functional connectivity. Yet, correlations are the end product of multiple cell interactions and, which do not always correspond to

\subsection*{Comparison with other uses of neural correlations}
\subsection*{Meaning of interactions}
\subsection*{Fitting probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
\subsection*{Future directions}

>>>>

\section*{Introduction}
\subsection*{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Sompolinsky:2001,Wilke:2002,Averbeck:2006,Josic:2009,Berens:2011,Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.  Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Arieli:1995,Chiu:2002,Kenet:2003,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.

@@

To illustrate challenges of estimating the correlation matrix from a finite sample, we briefly consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}. In covariance selection, the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero. We estimated the covariance matrix of a period of a somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (Fig.\;\ref{fig:01}\,A, B) and covariance selection (Fig.\;\ref{fig:01}\,C). Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales. The apparent similarity of the two estimates belies the utterly different partial correlation structure (Fig.\;\ref{fig:01}\,D, E): the regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients. Estimation by covariance selection is attractive for several reasons: First, it has fewer free parameters and is therefore less susceptible to sampling noise, yielding estimates that are, on average, closer to truth $\Sigma$. Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

@@

\section*{Discussion}
\subsection*{Neural correlations as indicators of functional connectivity}
Neural correlations have a rich history in neuroscience as indicators of \emph{functional connectivity}. When measured on temporal scales of a few milliseconds, peaks in the cross-correlations between pairs of neurons at latencies typical of synaptic transmission are often interpreted as evidence of direct synaptic connections between pairs of cells \cite{Moore:1970,Alonso:1998,Denman:2013}. Without targeted stimulations of specific cells or measurements of intracellular signals, such analysis remains the best test for probable monosynaptic connections in vivo. At longer temporal scales, neural correlations reflect activations involving large ensembles of neurons. Such correlations may have emergent structure with little  correspondence to synaptic connectivity, which  may better reflect computations performed by the circuit.  For this reason, firing rate correlations at scales of tens or hundreds milliseconds have been of interest for their implications on coding under the assumption of rate codes and linear downstream decoder \cite{Averbeck:2006}. The terms \emph{functional connectivity} is still often applied to slow correlations, similarly to its use in PET, fMRI, EEG, and MEG brain imaging studies, as an expression of task-related, stimulus-evoked, or spontaneous synchronization or correlation between brain regions, circuits, or cells without implying a mechanistic interpretation.

Calcium imaging allows simultaneous measurements of the activity of large populations of cells, even of entire nervous systems \cite{Leung:2013,Ahrens:2013}, with cell-wise resolution, but its temporal bandwidth is limited to the scale of tens or hundreds milliseconds by the kinetics of calcium-sensitive dyes. This limits their use to the analysis of slower correlations.

\subsection*{Correlation `structure'}
The term `structure' implies a reduced, simplified representation expressing the essential organization.  The structure of slow correlations is often expressed through low-rank approximations extracted by principal component analysis or factor analysis. These components show synchronous multineuronal activation patterns without revealing graphs of interactions between neurons.

Another reduced representation of correlation matrices is obtained by sparsifying its inverse, resulting in a graph of partial correlations. Under the assumption of a multivariate normal distribution or, more generally, an elliptical distribution, sparse partial correlations suggest a sparse matrix of dependencies, a probabilistic graphical model. This property has led some investigators to suggest that, in brain imaging studies, such approximation must serve as better proxies of direct interactions between brain regions than correlations \cite{Varoquaux:2012,Ryali:2012}.

\subsection*{Interpretation of functional connectivity from population calcium signals}
Nonetheless, some investigators have examined graphs of the most significant correlations in population calcium activity as proxy of connectivity \cite{Malmersjo:2013}.

While synaptic connectivity constitutes a directed weighted graph, most meaningful representation of functional organization at longer temporal scales are less certain. A common approach has been to examine the low-rank structure of the joint population activity extracted through principal component analysis, factor analysis, independent component analysis, linear embedding, and others. Functional connectivity is then represented as the eigenvectors, principle components, or factor loadings corresponding to the extracted components of the joint activity.

If only most significant components are retained, these two representations turn into, respectively, low-rank approximations and sparse inverse approximations. Low-rank representation suggest interpretation as common input onto large subsets of neurons or highly synchronous activations in the entire recorded population.

We sought to uncover the optimal low-dimensional approximation of the covariance matrix for efficient estimation of the true covariance matrices in cortical microcircuits on the temporal scale of 100--200 ms corresponding to the temporal resolution of spike rates from somatic calcium signals \cite{Cotton:2013}.  We considered sparse-inverse approximations, low-rank approximations, and mixtures of low-rank components with sparse partial correlations. We used cross-validation to determine which of such representations was closest to the true correlation matrix. As we illustrated by simulation, none of the estimators was intrinsically more efficient than the others and their relative performance was dependent on the structure of dependencies in the empirical data. We found that sparse inverse approximation performed better than low-rank approximations (Fig.\;S1\;E) and approximations by combinations of sparse and low-rank components consistently outperformed the other estimators. This finding is specific to the circuit we examined and may not generalize to other circuits or to less dense sampling of cells in electrophysiological recordings.

\subsection*{Probabilistic models and their interpretation}
On the millisecond scale, most common analyses of the functional connectivity have relied on cross-correlation analysis in attempts to detect signatures of direct synaptic connectivity.

Increasingly, the functional organization of neural circuits is inferred by fitting probabilistic models such as various forms of multivariate Gaussian models, linear-nonlinear models, Generalized Linear Models \cite{Pillow:2008}, and maximum entropy models \cite{Schneidman:2006,Tkacik:2006,Tang:2008,Shlens:2009}. In probabilistic models, functional dependencies correspond to multiplicative terms involving pairs or larger groups of neurons. If two neurons do not both participate in any of the multiplicative terms in the model, they are conditionally independent: the activity of one can be predicted from other neurons just equally well with or without considering the second neuron, suggesting a lack of direct interaction between them. Additionally, various probabilistic models allow inferring latent variables to explain joint patterns of in large groups of cells.  The uncovered networks of dependencies from such statistical models may arise from mechanistic interactions such as direct monosynaptic connections or from complex emergent processes involving large ensembles of cells. Because statistical models of the joint activity of populations of neurons use the activity of the entire population to infer functional dependencies, they to produce more accurate and relevant representations of the functional organization of the data than available from equivalent numbers of paired recordings.

However, no empirical evidence yet exists to demonstrate that statistical models of population activity reveal networks of dependences more informative than the conventional correlations with respect to the structural architecture of specific neural circuits.

\subsection*{Model selection to uncover structure best supported by data}
Model selection from various several probabilistic models

\subsection*{Functional connectivity from big neural data}
Transformational breakthroughs in systems neuroscience are anticipated to follow the ongoing technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far \cite{Alivisatos:2013}. It is far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques.

Such probabilistic models of population activity feature large numbers of parameters, growing quadratically or faster with population size. Estimation of large number of parameters is sensitive to sampling noise. To solve this problem, the recording durations must increase at least proportionally to the number of parameters or the model must be constrained by regularization.

\subsection*{Comparison with other uses of neural correlations}
\subsection*{Meaning of interactions}
\subsection*{Fitting and evaluation of probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
\subsection*{Future directions}
